Linear Ordered Attribute Grammar scheduling using SAT-solving

### 2014-10-15 paper/paper.tex
Subsection on chordal graph construction

Our algorithm has the following steps:
\begin{enumerate}
\item Construct a nonterminal dependency graph for each nonterminal
\item Construct a production dependency graph for each production
\item Make all graphs chordal
\item Introduce a SAT variable for each edge in any of the graphs, where edges from a production dependency graph and the corresponding nonterminal dependency graph are shared, so are represented by the same variable
\item Set the value of all variables corresponding to direct dependencies
\item Exclude all cycles of length three
\item Optionally add extra constraints for optimisations
\end{enumerate}
The first two steps have been explained in the previous sections. Step 3 is explained below\Thomas{Step 4 is missing}, and step 5 and 6 follow from that. Finally step 7 is explained in \sectref{sect:optimisations}.

@@

\Jeroen{todo: say something about heuristics for making the graph chordal and how many triangles there are}

>>>>

Our algorithm has the following steps:
\begin{enumerate}
\item Construct a nonterminal dependency graph for each nonterminal
\item Construct a production dependency graph for each production
\item Make all graphs chordal
\item Introduce a SAT variable for each edge in any of the graphs, where edges from a production dependency graph and the corresponding nonterminal dependency graph are shared, so are represented by the same variable
\item Set the value of all variables corresponding to direct dependencies
\item Exclude all cycles of length three
\item Optionally add extra constraints for optimisations
\end{enumerate}
The first two steps have been explained in the previous sections. Step 3 is explained below, step 4 is trivial, and step 5 and 6 follow from the explanation below. Finally step 7 is explained in \sectref{sect:optimisations}.

@@

\subsection{Chordal graph construction}
There are several algorithms for making a graph chordal. We use an algorithm based on the following alternative definition of a chordal graph:

\begin{definition}
An undirected graph is \emph{chordal} if and only if it has a 
perfect elimination order. A perfect elimination order is an
ordering $v_1,\ldots,v_n$ of the vertices of $G$ such that in the graph
$G[v_1,\ldots,v_i]$, $\univ{i}{1 \leq i \leq n}$, the vertex $v_i$ is 
\toindex{simplicial}. A vertex $v$ is called simplicial in a graph $G$ if
the neighbourhood of $v$ forms a connected component in $G$.
The graph $G[v_1,\ldots,v_i]$ is the induced subgraph of $G$ containing
only the vertices $v_i,\ldots,v_i$ and the edges between these vertices.
\end{definition}

From this definition we can construct the following algorithm for making a graph chordal:
\begin{enumerate}
 \item While the graph still contains vertices:
 \begin{enumerate}
     \item Select a vertex $v$ from the graph
     \item For every pair $(a,b)$ of unconnected vertices in the neighbourhood
        of $v$: 
        \begin{enumerate}
            \item Add the edge $(a\leftrightarrow b)$ to the graph
        \end{enumerate}
     \item Remove $v$, and all edges connected to $v$, from the graph.
 \end{enumerate}
\end{enumerate}

One important open question in this algorithm is the order in which the vertices should be chosen. In \sectref{sect:results} we show the results for several heuristics that we have implemented and tried on the UHC. We would like to remark that regardless of the heuristic used, this approach always leads to much smaller SAT problems than encoding transitivity in the SAT problem for ruling out cycles.
